A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-08-20 Papers/Paper/manuscript.Rmd
paper edits

We tweak the A-maze task to accomodate these longer passages and then have participants read the passages in the Maze. We compare our A-maze results with SPR data collected on this corpus by @futrellNaturalStoriesCorpus2020. We find that participants were able to read and understand these long passages using A-maze, and their RT profiles show a similar pattern to SPR but with less noise and spillover

@@

Many participants completed the Maze task with a high degree of accuracy and also answered the comprehension questions correctly.

@@

Participant accuracy reflects both how well participants can navigate the task and what quality the auto-generated distractors are. We calculated the per-word error rate for each participant and graphed it against their average reaction time. (To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy. This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story. These careful participants took around 1 second for each word selection, which is much slower than in eye-tracking or SPR. Another cluster of participants (in red) sped through the task, seemingly clicking randomly. This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.

@@

@@

TODO remind why we care so much about surprisal

We fitted generalized additive models to test whether the RTs from the Maze experiment showed a linear relationship with surprisal and whether the effect was limited the current word or had spilled over from the prior word. For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded. (See Methods for details on exclusions.)

@@

The smooths for the current and previous words surprisals are shown in Figure \@ref(fig:gam). Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal. All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data. All of the models show a flatter relationship between previous word surprisal and RT. This is a sign of localization as the previous word's surprisal is not affecting RT much. The linear relationship matches that found with other methodologies.

Given that the GAM models show a roughly linear relationship, we fit mixed linear models to quantify the influence of surprisal, frequency and word length. 
We built linear models with surprisal, frequency, length and surprisal x length and frequency x length effects from the current and previous words as predictors. We centered, but did not rescale, the predictors.

@@

As we can see in Table \@ref(tab:pre-error), we find large effects of surprisal and length, but minimal effects of frequency. The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019. Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word.

@@

@@

## maze works

## strong effects of suprisal, mostly localized

## better than SPR

## correlated w SPR, probably the same effects, just stronger & less noisy

## Why you care - broad applications that might want this method

## shill for error correction
We introduced a tweak to the paradigm for displaying Maze tasks that makes it useable for multi-sentence materials. In this error-correction Maze paradigm, participants read all the words because they can correct their mistakes and move on. We tested this method on the Natural Stories corpus, showing that, despite the oddness of the task, participants can read and understand a 1000 word story in this method. We additionally showed that the RTs generated in the Maze task show a linear relationship between the RT of a word and it's surprisal, but no relationship with the surprisal of the previous word. This provides additional evidence for the argument that the Maze task forces very incremental processing [@forsterMazeTaskMeasuring2009].

Actually the potential applicability and value is much broader, for any question where nailing down the locus on incremental processing difficulty is important, little spillover is very valuable.

This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty and thus benefits from the lack of spillover. 
It shows broadly the same effects as other methods; however the differences in scale of suprisal effects and the lack of frequency effects detected here are a reason to explore more. Comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].

While the error-correction paradigm is crucial to running long materials, it also provides some benefits even on shorter materials when using A-maze with variable participant populations. The error-correction compensates for some of the shortcomings of A-maze identified in @boyceMazeMadeEasy2020; poor distractors might still cause participants to make unavoidable mistakes, but they will still see the rest of the sentence, which may reduce frustration. One questions that remains is whether this post-mistake data is useable, which depends on whether RTs from a few words after a mistake show any differences from RTs before any mistakes. Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is a  hard-to-assess question of potential interest.

Even if post-mistake data is not analysed, it can be used to distinguish between errors due to inattentive participants or merely specific bad distractors early in the sentence. Researchers can calculate a per-word error rate for each participant; high per-word error rates are consistent with guessing, low error rates with errors clustered early the sentence are consistent with poor distractors. This per-word error rate metric measures participants task accuracy and provides a convenient and clear-cut way of controlling data quality after the fact.

@@

Between the distractor auto-generation process introduced in @boyceMazeMadeEasy2020 and the error-correction paradigm introduced here, the Maze paradigm is now easy to use on a wide range of materials. The ease of use, large effects, and forced incrementality make Maze a good complement to existing incremental processing methods. We encourage researchers to consider Maze as an option for doing incremental processing work. 
\newpage

>>>>

We tweak the A-maze task to accomodate these longer passages and then have participants read the passages in the Maze. We compare our A-maze results with SPR data collected on this corpus by @futrellNaturalStoriesCorpus2020. We find that participants were able to read and understand these long passages using A-maze, and the A-maze RT profiles show similar relationships to surprisal, word length, and word frequency as those seen with SPR, but with less noise and spillover.

@@

We first looked at participants overall performance on the Maze task and comrehension questions, in order to see if participants were able to read and understand a story in the Maze task. Many participants were able to complete the Maze task with a high degree of accuracy and answer the comprehension questions correctly.

@@

Participant accuracy, or how often participants choose the correct word over the distractor, reflects both the quality of the distractors and the focus and skill of the participant. We calculated the per-word error rate for each participant and graphed it against their average reaction time. (To avoid biasing the average if a participant took a pause before returning to the task, RTs greater than 5 seconds were excluded; this excluded `r nrow(data_long)` words, or `r nrow(data_long)/nrow(data_no_prac)*100 %>% round(2)`% of trials.) As seen in Figure \@ref(fig:errors)A, one cluster of participants (marked in green) made relatively few errors, with some reaching 99% accuracy. This confirms that the distractors were generally appropriate and shows that some participants maintained focus on the task for the whole story. These careful participants took around 1 second for each word selection, which is much slower than in eye-tracking or SPR. Another cluster of participants (in red) sped through the task, seemingly clicking randomly. This bimodal distribution is likely due to the mix of workers on Mechanical Turk, as we did not use qualification cutoffs.

@@

TODO: question: Does the current order of all Maze then all SPR results make sense? Or would be it better to have Maze-GAM, SPR-GAM, Maze-LM, SPR-LM, etc?

@@

Given that the Maze task worked, we next consider the relationship between surprisals and Maze RTs. Surprisal, a measure of overall word predictability in context, is linearly related to RT in eye-tracking and SPR. If Maze is measuring the same language processes, we would expect to see a linear relationship in the Maze task. Additionally, surprisal can be used to assess spillover, but comparing the predictive value of current and past word surprisal on current word reading time.

To assess the shape of the RT-surprisal relationship, we fit generalized additive models (GAMs). For these models, we only included data that occurred before any mistakes in the sentence; due to limits of model vocabulary, words with punctuation and some uncommon or proper nouns were excluded. We use surprisals generated by 4 different language models for robustness. (See Methods for details on language models, exclusions, and model fit.)

@@

The GAMs predicted relationship between current and previous word surprisals are shown in Figure \@ref(fig:gam); this is the main effect of either current or previous surprisal for an average length and frequency word, if the other surprisal value is 0 (TODO Not sure this is the correct interpretation! -- https://rdrr.io/cran/tidymv/man/get_gam_predictions.html is the function used for prediction). Note that for each of the models, high-surprisal words are rare, with much of the data for words between 0 and 15 bits of surprisal. All of the models show a roughly linear relationship between current word surprisal and RT, especially in the region with more data. All of the models show a flatter relationship between previous word surprisal and RT. This is a sign of localization as the previous word's surprisal is not affecting RT much. The linear relationship matches that found in SPR and eye-tracking. TODO comment on freq, length effects / non-linearities!

Given that the GAM models show a roughly linear relationship between RT and surprisal, we are justified in using surprisal as a predictor in a linear model to quantify its effect. In addition to surprisal, we also include the commonly studied effects of frequency and length as well as surprisal x length and frequency x length interactions. For all of these, we include the predictors for the current and previous word, and we centered, but did not rescale, all predictors. (See Methods for more details on these predictors and model fit process.)

@@

As shown in Table \@ref(tab:pre-error), we found large effects of surprisal and length, but minimal effects of frequency. The lack of frequency effects is somewhat surprising, but consistent with @shainLargescaleStudyEffects2019. Notably the coefficients for the lagged terms are small relative to the effects of surprisal and length of the current word, an indication that spillover is limited and effects are strongly localized.

@@

TODO mention at sentence level to control for spillover profiles and so single words are too noisy

Add interpretation

@@

We introduce error-correction Maze, a tweak on the presentation of Maze materials that makes Maze feasible for multi-sentence passages. We used A-maze distractors and this error-correction presentation to gather data on participants reading the Natural Stories in the Maze. We found that participants could read and comprehend the 1000 word stories, despite the slowness and added overhead of reading in the Maze task. This expands the domain of materials used with Maze beyond targeted single-sentence items to longer, naturalistic texts with sentence-to-sentence coherency.

We took advantage of the pre-existing SPR corpus on Natural Stories to compare the RT profiles between Maze and SPR. While these are not completely comparably as the subject populations differed, we are able to get a sense of the relationship between SPR and Maze results. Maze and SPR pick up on similar features in words, as shown by the correlations between Maze and SPR, and the fact that surprisal and length are predictive of RTs in both. However, the effect sizes of RT and surprisal are much larger is Maze, and the effects are more strongly localized on a single word. Further comparisons between different processing on the same materials could be useful for identifying how task demands influence language processing [ex. @bartekSearchOnlineLocality2011].

We find good localization on Maze, but it's worth noting that it's not absolute. Some of the models show small but statistically reliable relationships with surprisal from a past word. On the whole, however, these results support the idea that Maze forces language processing to be close to word-by-word, and that it's a relatively safe assumption that the RT of a word primarily reflects its own properties and not those of earlier words.

This extreme incrementality makes the Maze task a good target for any question that requires precisely determining the locus of incremental processing difficulty.

## Remarks about error-correction maze

With that in mind, we have a few remarks on the benefits of error-correction Maze and some avenues for further work on this method. Error-correction Maze makes for a smoother process. It dampens the effect of rare poor A-maze distractors; a distractor that fits the context might still cause participants to make an unavoidable mistake, but the mistake is less disruptive, as they still see the rest of the sentence. An open question for researchers is whether data from after a participant makes a mistake is useable, that is, does it show the same profile as pre-errors words, or are there traces from recovering from the mistake? Whether post-mistake data is high-quality and trustworthy enough to be included in analyses is a hard-to-assess question of potential interest.

Even if post-mistake data is not included in analyses, having this complete information can distinguish between errors due to inattentive participants and errors from specific bad distractors. Researchers can calculate a per-word error rate for each participant; high per-word error rates are consistent with guessing, low error rates with errors clustered early the sentence are consistent with poor distractors. This per-word error rate metric measures participants task accuracy and provides a convenient and clear-cut way of controlling data quality after the fact.

@@

Between error-correcting Maze and other innovations on the Maze paradigm such as A-maze and I-maze, the Maze task is versatile and can be used or adapted for a wide range of materials and questions of interest. We encourage researchers to use Maze as an incremental processing method, alone or in comparison with other methods.

\newpage
